\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-06-06 trunk/final-version/intro.tex
updates to final version

Autonomous platforms in the marine environment have had a substantial
and immediate impact for both civil and military
applications. Autonomous Underwater Vehicles (AUVs) have not only
altered the ability of oceanographers to reach beyond the surface and
make high resolution observations, but have impacted the engineering
methodologies behind sampling, control and robotics itself
\cite{Brierley08032002,ryan05,Thomas06,Yoerger01012007,Incze2009,Rigby10}. Yet
challenges remain, principally in making such robotic devices more
attune to the environment, to sample large swaths at the Meso-scale
($> 50$ km\textsuperscript{2}) and doing so systematically and
adaptively. Further, over the last two decades substantial
improvements in hardware including propulsion and battery technologies
have outpaced software especially in algorithmic methods in sampling
and control and in systematic software engineering development. Large
scale high resolution observations have become increasingly tractable
using autonomous platforms both on and below the surface.  This, in
turn, has lead to periodic calls for sustained exploration and
sampling \cite{rudnick03}, illustrating the high scientific value
placed on such data.  The complexity of oceanic processes and the
chronic under-sampling of the coastal oceans, especially of dynamic
features studied over a variety of spatial and temporal scales, has
made it challenging for robotic platforms to autonomously track and
sample. A prominent example of such a process is coastal algal blooms
which are patchy and could cover large coastal zones. Persistent
observation of such dynamic events dictates that our robotic assets
track and sample such patches which can evolve rapidly due to inherent
bio-geochemical activity, as well as advection and diffusion of the
water mass it resides in.

At the Monterey Bay Aquarium Research Institute (MBARI), a unique
inter-disciplinary collaboration between biologists, ecologists,
geneticists and roboticists on the \can (Controlled Agile Novel
Observation Network) program \cite{canon} has driven autonomous system
exploration and control. Fundamental questions related to sampling
strategies have driven the use cases for advanced concepts in
autonomy, specifically to Artificial Intelligence (AI) Planning and
Execution. Current work is also looking at how other fields in Machine
Intelligence especially Machine Learning can be leveraged to impact
the science of engineering systems in marine robotics and as a
consequence oceanography itself.

>>>>

Autonomous platforms in the marine environment have had a substantial
and immediate impact for both civil and military
applications. Autonomous Underwater Vehicles (AUVs) have not only
altered the ability of oceanographers to reach beyond the surface and
make high resolution observations, but have impacted the engineering
methodologies behind sampling, control and robotics itself
\cite{Brierley08032002,ryan05,Thomas06,Yoerger01012007,Incze2009,Rigby10}. Yet
challenges remain, principally in making such robotic devices more
attuned to the environment, to sample large swaths at the Meso-scale
($> 50$ km\textsuperscript{2}) and doing so systematically and
adaptively. Further, over the last two decades substantial
improvements in hardware including propulsion and battery technologies
have outpaced software especially in algorithmic methods in sampling
and control and in systematic software engineering development. Large
scale high resolution observations have become increasingly tractable
using autonomous platforms both on and below the surface.  This, in
turn, has lead to periodic calls for sustained exploration and
sampling \cite{rudnick03}, illustrating the high scientific value
placed on such data.  The complexity of oceanic processes and the
chronic under-sampling of the coastal oceans, especially of dynamic
features studied over a variety of spatial and temporal scales, has
made it challenging for robotic platforms to autonomously track and
sample. A prominent example of such a process is coastal algal blooms
which are patchy and could cover large coastal zones. Persistent
observation of such dynamic events dictates that our robotic assets
track and sample such patches which can evolve rapidly due to inherent
bio-geochemical activity, as well as advection and diffusion of the
water mass it resides in.

At the Monterey Bay Aquarium Research Institute (MBARI), a unique
inter-disciplinary collaboration between biologists, ecologists,
geneticists and roboticists on the \can (Controlled Agile Novel
Observation Network) program \cite{canon} has driven autonomous system
exploration and control. Fundamental questions related to sampling
strategies have driven the use cases for advanced concepts in
autonomy, specifically to Artificial Intelligence (AI) Planning and
Execution. Current work is also looking at how other fields in Machine
Intelligence, especially Machine Learning can be leveraged to impact
the science of engineering systems in marine robotics and as a
consequence oceanography itself.

### 2012-06-06 trunk/final-version/related.tex
Final Version to be sent to Publisher
updates to final version

Such techniques have proved their mettle in the first round of use of
AUVs; they've provided operators a simple way to handle control
complexity of the vehicle while returned data at far higher resolution
than would have been possible using traditional ship-based methods and
at substantially reduced amortized costs of platform operation and
ownership over multiple years. AUV operators and clients for their
data products have reasons to be well satisfied with results thus
far. However, as hardware and more cost-effective sensors become more
robust to the environment and as mission durations increase (as
demonstrated by glider experiments such as \cite{rucool11}), sustained
presence in the ocean requires the ability to deal with off-nominal
conditions, that balances the needs of high-level mission management,
low-level navigation, instrument control and detection of unstructured
and poorly understood phenomena. Earlier reactive methods are unable
to perform such a balancing act without skewering the foundations of
the software development methodology leading to code bloat or worse.
\cite{carreras06} provides a good overview of AUV control
architectures essentially based on reactive control methods.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[scale=0.4]{figs/planner-arch.pdf}
 \caption{\small A general architectural block diagram for an AI based
   Planner.}
 \label{fig:planner}
\end{figure}

>>>>

Such techniques have proved their mettle in the first round of use of
AUVs; they have provided operators a simple way to handle control
complexity of the vehicle while returned data at far higher resolution
than would have been possible using traditional ship-based methods and
at substantially reduced amortized costs of platform operation and
ownership over multiple years. AUV operators and clients for their
data products have reasons to be well satisfied with results thus
far. However, as hardware and more cost-effective sensors become more
robust to the environment and as mission durations increase (as
demonstrated by glider experiments such as \cite{rucool11}), sustained
presence in the ocean requires the ability to deal with off-nominal
conditions, that balances the needs of high-level mission management,
low-level navigation, instrument control and detection of unstructured
and poorly understood phenomena. Earlier reactive methods are unable
to perform such a balancing act without skewering the foundations of
the software development methodology leading to code bloat or worse.
\cite{carreras06} provides a good overview of AUV control
architectures essentially based on reactive control methods.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[scale=2]{figs/planner-arch.jpeg}
 \caption{\small A general architectural block diagram for an AI based
   Planner.}
 \label{fig:planner}
\end{figure}
